---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# osmexport

<!-- badges: start -->
<!-- badges: end -->

**osmexport** is an R package useful to download, read, prepare and (rudimentarily) visualise OpenStreetMap (OSM) exports like the `map.osm` file you would get from the [OSM website's "Export" page](https://www.openstreetmap.org/export).

This package might not be what you're after: please see below the section "Other OSM-related R packages" to find more powerful tools.

## Installation

You can install the development version of osmexport from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stragu/osmexport")
```

## Available functions

* `oexp_download()`: download an OSM export by defining its bounding box, save it to file.
* `oexp_read()`: read a `.osm` file as an object of class `osm`
* `oexp_separate_tags()`: separate the tags contained in the `other_tags` column. This is done by default when importing the data with `oexp_read()`

## The `osm` class

The `osm` S3 class – very much subject to changing in name and properties – has `print()` and `plot()` methods. It is made of [`sf`](https://r-spatial.github.io/sf/) ("simple features") spatial objects.

## Example

The functions available are designed to be pipeable. For example, to download and read the area around [Te Kura Tatauranga, Waipapa Taumata Rau](https://www.auckland.ac.nz/en/science/about-the-faculty/department-of-statistics.html):

```{r Te_Kura_Tatauranga}
library(osmexport)
TKT <- c(174.76598, -36.85440, 174.77019, -36.85129) |>
  oexp_download() |>
  oexp_read()
# see what the object contains
TKT
```

We can now use the default plot method to have a glimpse at the data:

```{r plot}
plot(TKT)
```

Because the parts of the object are of class `sf`, they can be processed with sf and well as dplyr functions:

```{r process_data, message=FALSE}
library(dplyr)
library(sf)
# most common "building" values in polygons
TKT$multipolygons |>
  st_drop_geometry() |>         # remove geometry column
  filter(!is.na(building)) |>   # only keep buildings
  count(building, sort = TRUE)  # most commons values at the top
```

## Contributions and feedback

Contributions are welcome and appreciated. You can contribute to this package by:

* Testing it and [reporting issues](https://github.com/stragu/osmexport/issues) you encounter
* Suggesting a change with a pull request (but please discuss your idea in an issue beforehand)
* Writing a new vignette with an interesting worked example

Note that all contributions to the codebase will be released under the [GPL](LICENSE.md), and submitting your contribution to this repository is an implicit agreement.

We expect contributors to respect UQRUG's [Code of Conduct](https://gitlab.com/stragu/uqrug/blob/master/Code_of_Conduct.md).

If a conversation or comment does not belong in a public issue report, please contact the maintainer listed in the [package description](DESCRIPTION).

## Data licence

OSM data included in this package (in the `./inst` directory) and downloaded with the `oexp_download()` function is © OpenStreetMap contributors, and the conditions of its reuse are defined by the [ODbL licence](https://opendatacommons.org/licenses/odbl/). You likely need to include this information on anything derived from it. Find out more on the [OSM website](https://www.openstreetmap.org/copyright).

## Other OSM-related R packages

There are many other OSM-related R packages that might be more suitable for what you are hoping to achieve. Notable packages available on CRAN are:

* [osmadata](https://cran.r-project.org/web/packages/osmdata) uses the Overpass API to download datasets based on location and tags. Use this for a more targeted download.
* [osmextract](https://cran.r-project.org/web/packages/osmextract) uses different providers to download chunks of large data dumps, for specific regions and dates. Some providers offer data already prepared for particular uses (e.g. data relevant to cycling).
* [osmplotr](https://cran.r-project.org/web/packages/osmplotr) downloads OSM data with the Overpass API and provides many tools to render the data.
* [osmar](https://cran.r-project.org/web/packages/osmar/index.html) accesses OSM data from different sources and converts it to various formats. Note that it hasn't been updated since 2013.

Transport-specific packages:

* [stplanr](https://cran.r-project.org/web/packages/stplanr) provides tools for sustainable transport planning, making use of the osrm package.
* [osrm](https://cran.r-project.org/web/packages/osrm) bridges R with the OSRM API (routing service based on OpenStreetMap data).
* [opentripplanner](https://docs.ropensci.org/opentripplanner/) provides an interface to OpenTripPlanner (OTP), a routing service that relies on OpenStreetMap data.

To acquire basemaps:

* [OpenStreetMap](https://cran.r-project.org/web/packages/OpenStreetMap) allows downloading rendered OSM data from various servers.
* [OSMscale](https://cran.r-project.org/web/packages/osmplotr) plots spatial data with OSM basemaps and corresponding scale bars.

Other interactive visualisation packages automatically fetch rendered OSM data as basemaps, like [leaflet](https://rstudio.github.io/leaflet/) and [tmap](https://r-tmap.github.io/tmap/).
